---
title: "knb-lter-cap.702"
output: html_document
---

# README

contents moved to README.md


# location helpers

```{r SNAZ_sites, eval=TRUE, echo=TRUE, message=TRUE}

snaz_sites <- c('KP', 'CB', 'PIE', 'MR', 'MS', 'SW', 'BV', 'ENC')
salt_sites <- c('Ave7th', 'centralNorth', 'centralSouth', 'Price')
```

# runoff chemistry

## runoff chemistry: query

```{r runoff-chemistry-table}

runoff_chemistry_query <- glue::glue_sql("
SELECT
  s.bottle,
  sites.abbreviation AS runoff_location,
  s.sample_datetime AS runoff_datetime,
  s.comments AS sample_comments,
  s.\"lab_pH\" AS pH,
  r.replicate,
  a.analysis_name,
  r.concentration AS analysis_concentration,
  dq.data_qualifier_label AS analysis_data_qualifier,
  r.comments AS analysis_comments
FROM
  stormwater.samples s
  LEFT JOIN stormwater.results r ON (r.sample_id = s.sample_id)
  JOIN stormwater.analysis a ON (r.analysis_id = a.analysis_id)
  JOIN stormwater.sites ON (s.site_id = sites.site_id)
  LEFT JOIN stormwater.data_qualifier dq ON (dq.data_qualifier_id = r.data_qualifier)
WHERE
  EXTRACT (YEAR FROM s.sample_datetime) > 2007 AND
  -- r.analysis_id IS NOT NULL AND
  (bottle NOT ILIKE '%rain%' OR bottle IS NULL) AND
  sites.abbreviation IN ({snaz_sites*})
;
",
.con = DBI::ANSI()
)

runoff_chemistry <- DBI::dbGetQuery(pg, runoff_chemistry_query)

runoff_chemistry <- runoff_chemistry |>
  dplyr::mutate(
    blank           = as.factor(ifelse(grepl("blk|blank", bottle, ignore.case = T), "TRUE", "FALSE")),
    runoff_location = as.factor(runoff_location),
    analysis_name   = as.factor(analysis_name)
    ) |>
  dplyr::select(
    runoff_location,
    runoff_datetime,
    blank,
    everything(),
    -bottle
    ) |>
  dplyr::arrange(
    runoff_location,
    analysis_name,
    runoff_datetime
  )

```

## runoff chemistry: table

```{r runoff-chemistry-table}

try({
  capeml::write_attributes(runoff_chemistry, overwrite = FALSE)
  capeml::write_factors(runoff_chemistry, overwrite = FALSE)
})

runoff_chemistry_desc <- "chemistry of runoff-generating storms at CAP LTER Stormwater Nitrogen in Arizona (SNAZ) project sampling sites"

runoff_chemistry_DT <- capeml::create_dataTable(
  dfname         = runoff_chemistry,
  description    = runoff_chemistry_desc,
  dateRangeField = "runoff_datetime",
  overwrite      = TRUE
)

```

# rainfall

## rainfall chemistry

```{r rainfall-chemistry}

rainfallChemistry <- DBI::dbGetQuery(pg, "
SELECT
  sites.abbreviation AS rain_location,
  s.sample_datetime AS rain_datetime,
  r.replicate,
  a.analysis_name AS analysis,
  r.concentration,
  dq.data_qualifier_label AS data_qualifier,
  r.comments
FROM
  stormwater.results r
  JOIN stormwater.samples s ON (s.sample_id = r.sample_id)
  JOIN stormwater.analysis a ON (r.analysis_id = a.analysis_id)
  JOIN stormwater.sites ON (s.site_id = sites.site_id)
  LEFT JOIN stormwater.data_qualifier dq ON (dq.data_qualifier_id = r.data_qualifier)
WHERE
  EXTRACT (YEAR FROM s.sample_datetime) > 2007 AND
  r.analysis_id IS NOT NULL AND
  bottle ILIKE '%rain%'
ORDER BY
  sites.abbreviation,
  s.sample_datetime,
  a.analysis_name
;
"
)

rainfall_chemistry <- rainfallChemistry |>
  dplyr::mutate(
    rain_location = as.factor(rain_location),
    analysis      = as.factor(analysis),
    rain_datetime = as.Date(rain_datetime)
  )

try({
  capeml::write_attributes(rainfall_chemistry, overwrite = FALSE)
  capeml::write_factors(rainfall_chemistry, overwrite = FALSE)
})

rainfall_chemistry_desc <- "chemistry of rainfall at CAP LTER Stormwater Nitrogen in Arizona (SNAZ) project sampling sites"

rainfall_chemistry_DT <- capeml::create_dataTable(
  dfname         = rainfall_chemistry,
  description    = rainfall_chemistry_desc,
  dateRangeField = "rain_datetime"
)

```

## rainfail amount

```{r rainfall}

rainfall <- DBI::dbGetQuery(pg, "
SELECT
  sites.abbreviation AS rain_location,
  r.event_datetime AS rain_datetime,
  r.rainfall_quantity AS rain_quantity
FROM
  stormwater.rainfall r
  JOIN stormwater.sites ON (r.site_id = sites.site_id)
  ;
  "
)

rainfall <- rainfall |>
  dplyr::mutate(
      rain_location = as.factor(rain_location)
    )

dplyr::glimpse(rainfall)

try({
  capeml::write_attributes(rainfall, overwrite = FALSE)
  capeml::write_factors(rainfall, overwrite = FALSE)
})

rainfall_desc <- "amount of precipitation as measured by tipping-bucket rain gauge at CAP LTER Stormwater Nitrogen in Arizona (SNAZ) project sampling sites"

rainfall_DT <- capeml::create_dataTable(
  dfname         = rainfall,
  description    = rainfall_desc,
  dateRangeField = "rain_datetime",
  overwrite      = TRUE
)

```

# discharge

```{r discharge}

snaz_q_query <- glue::glue_sql("
SELECT
  sites.abbreviation AS discharge_location,
  r.event_datetime AS discharge_datetime,
  r.water_height AS water_height,
  r.discharge AS raw_discharge,
  r.discharge_corrected AS edited_discharge
FROM
  stormwater.discharge r
  JOIN stormwater.sites ON (r.site_id = sites.site_id)
WHERE
  sites.abbreviation IN ({snaz_sites*})
;
",
.con = DBI::ANSI()
)

snaz_q <- DBI::dbGetQuery(pg, snaz_q_query)

lm_sgc_q_query <- glue::glue_sql("
SELECT
  sites.abbreviation AS discharge_location,
  r.event_datetime AS discharge_datetime,
  r.water_height AS water_height,
  r.discharge AS raw_discharge,
  r.discharge_corrected AS edited_discharge
FROM
  stormwater.discharge r
  JOIN stormwater.sites ON (r.site_id = sites.site_id)
WHERE
  sites.abbreviation IN ('SGC', 'LM') AND
  -- (r.discharge IS NOT NULL OR r.discharge_corrected IS NOT NULL)
  r.event_datetime <= '2012-08-24'
;
",
.con = DBI::ANSI()
)

lm_sgc_q <- DBI::dbGetQuery(pg, lm_sgc_q_query)

discharge <- dplyr::bind_rows(
    snaz_q,
    lm_sgc_q
  ) |>
  dplyr::arrange(
    discharge_location,
    discharge_datetime
    ) |>
  dplyr::mutate(
    discharge_location = as.factor(discharge_location)
  )

try({
  capeml::write_attributes(discharge, overwrite = FALSE)
  capeml::write_factors(discharge, overwrite = FALSE)
})

discharge_desc <- "discharge at CAP LTER Stormwater Nitrogen in Arizona (SNAZ) project sampling sites"

discharge_DT <- capeml::create_dataTable(
  dfname         = discharge,
  description    = discharge_desc,
  dateRangeField = "discharge_datetime"
)

```

# particulates

```{r particulates}

particulates_query <- glue::glue_sql("
SELECT
  s.bottle,
  sites.abbreviation AS runoff_location,
  s.sample_datetime AS runoff_datetime,
  r.replicate AS replicate,
  r.filter_initial AS filter_wt,
  r.filter_dry AS filter_dry_wt,
  r.volume_filtered AS vol_filtered,
  r.filter_ashed AS filter_ash_wt,
  r.data_qualifier AS data_qualifier,
  r.comments AS comments
FROM
  stormwater.solids r
  JOIN stormwater.samples s ON (s.sample_id = r.sample_id)
  JOIN stormwater.sites ON (s.site_id = sites.site_id)
  LEFT JOIN stormwater.data_qualifier dq ON (dq.data_qualifier_id = r.data_qualifier)
WHERE
  sites.abbreviation IN ({snaz_sites*})
;
",
.con = DBI::ANSI()
)

particulates <- DBI::dbGetQuery(pg, particulates_query)

particulates <- particulates |>
  dplyr::mutate(
    blank           = as.factor(ifelse(grepl("blk|blank", bottle, ignore.case = T), "TRUE", "FALSE")),
    runoff_location = as.factor(runoff_location),
    data_qualifier  = as.character(data_qualifier)
  ) |>
  dplyr::select(-bottle)

dplyr::glimpse(particulates)

try({
  capeml::write_attributes(particulates, overwrite = FALSE)
  capeml::write_factors(particulates, overwrite = FALSE)
})

particulates_desc <- "mass of particulates in stormwater during runoff-generating storms at CAP LTER Stormwater Nitrogen in Arizona (SNAZ) project sampling sites"

particulates_DT <- capeml::create_dataTable(
  dfname         = particulates,
  description    = particulates_desc,
  dateRangeField = "runoff_datetime"
)

```

# sampling_location

We include the LT sites LM, SBG, and IBW in the sampling location entity as
rainfall was collected and reported from those sites in this dataset. This is a
bit confusing as the runoff chemistry from those sites are reported in the LT
(knb-lter-cap.624) dataset.

```{r sampling-location, eval=TRUE}

sampling_location_query <- glue::glue_sql("
  SELECT
    abbreviation AS location_identifier,
    site_name AS location_description,
    latitude,
    longitude
  FROM
    stormwater.sites
  WHERE
    sites.abbreviation NOT IN ({salt_sites*})
  ;
  ",
  .con = DBI::ANSI()
)

sampling_location <- DBI::dbGetQuery(pg, sampling_location_query) |>
  sf::st_as_sf(
    coords = c("longitude", "latitude"),
    crs    = 4326
  )

try({
  capeml::write_attributes(sampling_location, overwrite = FALSE)
})

sampling_location_desc <- "CAP LTER Stormwater Nitrogen in Arizona (SNAZ) project sampling sites"

sampling_location_SV <- capemlGIS::create_vector(
  vector_name = sampling_location,
  description = sampling_location_desc
)

```

# analyses

```{r analyses}

analytes <- DBI::dbGetQuery(pg,"
SELECT
  DISTINCT ON (analysis_name)
  analysis_name as analysis,
  analysis_description AS description,
  units,
  instrument
FROM
  stormwater.analysis
  RIGHT JOIN stormwater.results r ON (r.analysis_id = analysis.analysis_id)
WHERE
  analysis_name IS NOT NULL
;
")

dplyr::glimpse(analytes)

try({
  capeml::write_attributes(analytes, overwrite = FALSE)
})

analytes_desc <- "catalog and details of water chemistry analytes measured as part of CAP LTER stormwater monitoring"

analytes_DT <- capeml::create_dataTable(
  dfname      = analytes,
  description = analytes_desc
)

```

# indian_bend_wash_catchments

Here we keep the source file (stormwater_catchments.geojson) in the repository
so that we can easily reconstruct indian_bend_wash_catchments when updating.

```{r indian-bend-wash-catchments, eval=TRUE}

# sf::st_write(
#   obj        = indian_bend_wash_catchments,
#   dsn        = "stormwater_catchments.geojson",
#   driver     = "geojson",
#   delete_dsn = TRUE
# )

indian_bend_wash_catchments <- sf::read_sf("stormwater_catchments.geojson")

try({
  capeml::write_attributes(indian_bend_wash_catchments, overwrite = FALSE)
})

indian_bend_wash_catchments_desc <- "The location and configuration of watersheds that are or have been sampled as part of CAP LTER Stormwater Nitrogen in Arizona (SNAZ) project research on stormwater dynamics in the greater Phoenix metropolitan area."

indian_bend_wash_catchments_SV <- capemlGIS::create_vector(
  vector_name = indian_bend_wash_catchments,
  description = indian_bend_wash_catchments_desc
)

```

# methods

```{r methods}

main <- list(description = capeml::read_markdown("methods.md"))

# provenance: long-term stormwater monitoring
stormwater_lt            <- emld::as_emld(EDIutils::get_provenance_metadata("knb-lter-cap.624.4"))
stormwater_lt$`@context` <- NULL
stormwater_lt$`@type`    <- NULL

enhancedMethods <- EML::eml$methods(methodStep = list(main, stormwater_lt))

```

# people

```{r people}

# creators

dan <- gioseml::create_role(
  firstName = "dan",
  lastName  = "childers",
  roleType  = "creator"
)

stevan <- gioseml::create_role(
  firstName = "ste",
  lastName  = "earl",
  roleType  = "creator"
)

nancy <- gioseml::create_role(
  firstName = "nan",
  lastName  = "grimm",
  roleType  = "creator"
)

rebecca <- gioseml::create_role(
  firstName = "rebe",
  lastName  = "hale",
  roleType  = "creator"
)

laura <- gioseml::create_role(
  firstName = "laura",
  lastName  = "turnb",
  roleType  = "creator"
)

creators <- list(
  dan,
  stevan,
  nancy,
  rebecca,
  laura
)


# metadata providers

stevan <- gioseml::create_role(
  firstName = "ste",
  lastName  = "earl",
  roleType  = "meta"
)

quincy <- gioseml::create_role(
  firstName = "qui",
  lastName  = "stewar",
  roleType  = "meta"
)

sally <- gioseml::create_role(
  firstName = "sally",
  lastName  = "wittlin",
  roleType  = "meta"
)

metadataProvider <- list(
  stevan,
  quincy,
  sally
)

```

# coverages

```{r coverages}

begindate <- format(min(runoff_chemistry$runoff_datetime), "%Y-%m-%d")
enddate   <- format(max(runoff_chemistry$runoff_datetime), "%Y-%m-%d")
geo_desc  <- yaml::yaml.load_file("config.yaml")[["geographic_description"]]

coverage <- EML::set_coverage(
  begin                 = begindate,
  end                   = enddate,
  geographicDescription = geo_desc,
  west                  = sf::st_bbox(sampling_location)[["xmin"]],
  east                  = sf::st_bbox(sampling_location)[["xmax"]],
  north                 = sf::st_bbox(sampling_location)[["ymax"]],
  south                 = sf::st_bbox(sampling_location)[["ymin"]]
)

```

# literature cited

```{r literature-cited, eval=TRUE}

grimm_2004 <- capeml::create_citation("https://doi.org/10.1029/153GM11")
grimm_2005 <- capeml::create_citation("https://doi.org/10.1899/04-027.1")
halea      <- capeml::create_citation("https://doi.org/10.1021/es501039t")
haleb      <- capeml::create_citation("https://doi.org/10.1007/s10021-014-9812-2")
roach_2011 <- capeml::create_citation("https://doi.org/10.1890/10-1613.1")
roach_2008 <- capeml::create_citation("https://doi.org/10.1641/B580808")
walsh      <- capeml::create_citation("https://doi.org/10.1899/04-028.1")

citations <- list(
  citation = list(
    grimm_2004,
    grimm_2005,
    halea,
    haleb,
    roach_2011,
    roach_2008,
    walsh
  ) # close list of citations
) # close citation

```

# customUnits

Note that eval is set to `FALSE`; this is intentional as this unit is not
applied.

```{r custom-units, eval=FALSE}

custom_units <- rbind(
  data.frame(
    id             = "milligramPerKilogram",
    unitType       = "massPerMass",
    parentSI       = "gramsPerGram",
    multiplierToSI = 0.000001,
    description    = "millgram of element per kilogram of material"
  )
)

unitList <- EML::set_unitList(
  custom_units,
  as_metadata = TRUE
)

```

# dataset

Optionally, provide: scope, abstract, methods, keywords, publication date.
Projects scopes include lter (default), urex, ltreb, and som.

```{r construct-dataset}

dataset <- capeml::create_dataset()
```

# eml

```{r construct_eml, eval=TRUE}

eml <- capeml::create_eml()
```

```{r validate_eml, eval=TRUE}

EML::eml_validate(eml)
```

```{r eml_to_file, eval=TRUE}

# write the eml to file
capeml::write_cap_eml()
```

# file placement

```{r package-details, eval=TRUE}

# retrieve package details from config.yaml
if (!file.exists("config.yaml")) {
  stop("config.yaml not found")
}
package_id <- eml[["packageId"]]
identifier <- yaml::yaml.load_file("config.yaml")$identifier
```

```{r preview-data-file-to-upload}

# preview data set files that will be uploaded to S3
list.files(pattern = paste0(identifier, "_"))
```

Move data and final xml files to respective ASU locations.

```{r S3-helper-functions}

# functions and setting for uploading to S3
library(aws.s3)
source("~/Documents/localSettings/aws.s3")
```

```{r upload-data-S3}

# upload files to S3
lapply(list.files(pattern = paste0(identifier, "_")), gioseml::data_to_amz)
```


# EDI

## EDI: login

```{r edi-login, eval=TRUE, echo=TRUE, message=TRUE}

EDIutils::login(
  userId   = keyring::key_get("edi_user"),
  userPass = keyring::key_get("edi_pass")
)

```

## EDI: evaluate

```{r edi-evaluate, eval=TRUE, echo=TRUE, message=TRUE}

evaluation <- EDIutils::evaluate_data_package(
  eml         = paste0(eml[["packageId"]], ".xml"),
  useChecksum = FALSE,
  env         = "staging"
)

Sys.sleep(8)

eval_status <- EDIutils::check_status_evaluate(
  transaction = evaluation,
  env         = "staging"
)

if (eval_status) {

  # evalution summary

  EDIutils::read_evaluate_report_summary(
    transaction = evaluation,
    env         = "staging"
  )

}

# evalution detailed

# EDIutils::read_evaluate_report(
#   transaction = evaluation,
#   env         = "staging"
# )

```

## EDI: update

```{r edi-update, eval=TRUE, echo=TRUE, message=TRUE}

EDIutils::update_data_package(
  eml         = paste0(eml[["packageId"]], ".xml"),
  useChecksum = TRUE,
  env         = "production"
)

```

## EDI: logout

```{r edi-logout, eval=TRUE, echo=TRUE, message=TRUE}

EDIutils::logout()

```


```{r clean-up}

# remove data files
dataFilesToRemove <- dir(pattern = paste0(identifier, "_"))
file.remove(dataFilesToRemove)

# EML to S3
gioseml::eml_to_amz(list.files(pattern = "knb.+xml"))

# EML to cap-data-eml and remove file from project
file.copy(list.files(pattern = "knb.+xml"), "/home/srearl/localRepos/cap-metadata/cap-data-eml/")
file.remove(list.files(pattern = "knb.+xml"))
```

